---
title: "Data 607 Week 10 Assignment"
author: "Trishita Nath"
date: "4/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In Text Mining with R, Chapter 2 looks at Sentiment Analysis.  In this assignment, you should start by getting the primary example code from chapter 2 working in an R Markdown document.  You should provide a citation to this base code.  You’re then asked to extend the code in two ways:

* Work with a different corpus of your choosing, and 
* Incorporate at least one additional sentiment lexicon (possibly from another R package that you’ve found through research).

### Loading Libraries
```{r}
library(tidyverse)
library(tidytext)
library(textdata)
library(janeaustenr)
library(stringr)
library(wordcloud)
library(reshape2)
library(gutenbergr)
```

## Part 1: Example Code

### 1.1 The sentiments datasets

```{r echo=TRUE, eval=TRUE}
get_sentiments("afinn")
get_sentiments("bing")
get_sentiments("nrc")
```

### 1.2 Sentiment analysis with inner join

```{r echo=TRUE, eval=TRUE}
#words with a joy score from the NRC lexicon
tidy_books <- austen_books() %>%
  group_by(book) %>%
  mutate(
    linenumber = row_number(),
    chapter = cumsum(str_detect(text, 
                                regex("^chapter [\\divxlc]", 
                                      ignore_case = TRUE)))) %>%
  ungroup() %>%
  unnest_tokens(word, text)

# Now that the text is in a tidy format with one word per row, we are ready to do the sentiment analysis
nrc_joy <- get_sentiments("nrc") %>% 
  filter(sentiment == "joy")

tidy_books %>%
  filter(book == "Emma") %>%
  inner_join(nrc_joy) %>%
  count(word, sort = TRUE)

# Next, we count up how many positive and negative words there are in defined sections of each book
jane_austen_sentiment <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(book, index = linenumber %/% 80, sentiment) %>%
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>% 
  mutate(sentiment = positive - negative)

# Now we can plot these sentiment scores across the plot trajectory of each novel
ggplot(jane_austen_sentiment, aes(index, sentiment, fill = book)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~book, ncol = 2, scales = "free_x")
```

### 1.3 Comparing the three sentiment dictionaries

```{r echo=TRUE, eval=TRUE}
# Choose only the words from the one novel we are interested in
pride_prejudice <- tidy_books %>% 
  filter(book == "Pride & Prejudice")

pride_prejudice

# define larger sections of text that span multiple lines and find the net sentiment in each of these sections of text
afinn <- pride_prejudice %>% 
  inner_join(get_sentiments("afinn")) %>% 
  group_by(index = linenumber %/% 80) %>% 
  summarise(sentiment = sum(value)) %>% 
  mutate(method = "AFINN")

bing_and_nrc <- bind_rows(
  pride_prejudice %>% 
    inner_join(get_sentiments("bing")) %>%
    mutate(method = "Bing et al."),
  pride_prejudice %>% 
    inner_join(get_sentiments("nrc") %>% 
                 filter(sentiment %in% c("positive", 
                                         "negative"))
    ) %>%
    mutate(method = "NRC")) %>%
  count(method, index = linenumber %/% 80, sentiment) %>%
  pivot_wider(names_from = sentiment,
              values_from = n,
              values_fill = 0) %>% 
  mutate(sentiment = positive - negative)

# Bind them together and visualize them 
bind_rows(afinn, 
          bing_and_nrc) %>%
  ggplot(aes(index, sentiment, fill = method)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~method, ncol = 1, scales = "free_y")

# how many positive and negative words are in these lexicons
get_sentiments("nrc") %>% 
  filter(sentiment %in% c("positive", "negative")) %>% 
  count(sentiment)

get_sentiments("bing") %>% 
  count(sentiment)


```

### 1.4 Most common positive and negative words

```{r echo=TRUE, eval=TRUE}
bing_word_counts <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  ungroup()

bing_word_counts

# This can be shown visually, and we can pipe straight into ggplot2
bing_word_counts %>%
  group_by(sentiment) %>%
  slice_max(n, n = 10) %>% 
  ungroup() %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, scales = "free_y") +
  labs(x = "Contribution to sentiment",
       y = NULL)
# Anomaly for the word "miss"
custom_stop_words <- bind_rows(tibble(word = c("miss"),  
                                      lexicon = c("custom")), 
                               stop_words)

custom_stop_words
```

### 1.5 Wordclouds

```{r echo=TRUE, eval=TRUE}
tidy_books %>%
  anti_join(stop_words) %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 100))

#  sentiment analysis to tag positive and negative words using an inner join, then find the most common positive and negative words
tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  acast(word ~ sentiment, value.var = "n", fill = 0) %>%
  comparison.cloud(colors = c("gray20", "gray80"),
                   max.words = 100)


```

### 1.6 Looking at units beyond just words

```{r echo=TRUE, eval=TRUE}
p_and_p_sentences <- tibble(text = prideprejudice) %>% 
  unnest_tokens(sentence, text, token = "sentences")

p_and_p_sentences$sentence[2]

austen_chapters <- austen_books() %>%
  group_by(book) %>%
  unnest_tokens(chapter, text, token = "regex", 
                pattern = "Chapter|CHAPTER [\\dIVXLC]") %>%
  ungroup()

austen_chapters %>% 
  group_by(book) %>% 
  summarise(chapters = n())

bingnegative <- get_sentiments("bing") %>% 
  filter(sentiment == "negative")

wordcounts <- tidy_books %>%
  group_by(book, chapter) %>%
  summarize(words = n())

tidy_books %>%
  semi_join(bingnegative) %>%
  group_by(book, chapter) %>%
  summarize(negativewords = n()) %>%
  left_join(wordcounts, by = c("book", "chapter")) %>%
  mutate(ratio = negativewords/words) %>%
  filter(chapter != 0) %>%
  slice_max(ratio, n = 1) %>% 
  ungroup()
```

## Part 2: Extension - My Bondage and My Freedom by Frederick Douglass

I will analyze text “My Bondage and My Freedom” by Frederick Douglass. I will use the gutenbergr library to search and download it. 

### 2.1 The sentiments datasets

```{r echo=TRUE, eval=TRUE}
bondage_count <- gutenberg_download(202)
bondage_count
```
### 2.2 Tidy the data

```{r echo=TRUE, eval=TRUE}
#removing the first 763 rows of text which are table of contents
bondage_count <- bondage_count[c(763:nrow(bondage_count)),]

#using unnest_tokens to have each line be broken into indidual rows. 
bondage <- bondage_count %>% unnest_tokens(word, text)
bondage

bondage_index <- bondage_count %>% 
  filter(text != "") %>%
  mutate(linenumber = row_number(),
         chapter = cumsum(str_detect(text, regex("(?<=Chapter )([\\dII]{1,3})", ignore_case =  TRUE)))) 

bondage_index
```

### 2.3 Most frequent postive words

```{r echo=TRUE, eval=TRUE}
bondage %>% 
  inner_join(get_sentiments("bing")) %>% 
  filter(sentiment == "positive") %>%
  count(word, sentiment, sort = TRUE) %>% 
  top_n(10) %>%
  mutate(word = reorder(word, desc(n))) %>%
  ggplot() + 
  aes(x = word, y = n) +
  labs(title = "Most Frequent Positive Words") + 
  ylab("Count") + 
  xlab("Word") +
  geom_col() + 
  geom_text(aes(label = n, vjust = -.5)) + 
  theme(
    panel.background = element_rect(fill = "white", color = NA),
    axis.text.y = element_blank(), 
    axis.ticks.y = element_blank(),
    plot.title = element_text(hjust = 0.5)
  )
```

### 2.4 Most frequent negative words

```{r echo=TRUE, eval=TRUE}
bondage %>% 
  inner_join(get_sentiments("bing")) %>% 
  filter(sentiment == "negative") %>%
  count(word, sentiment, sort = TRUE) %>% 
  top_n(10) %>%
  mutate(word = reorder(word, desc(n))) %>%
  ggplot() + 
  aes(x = word, y = n) +
  labs(title = "Most Frequent Negative Words") + 
  ylab("Count") + 
  xlab("Word") +
  geom_col() + 
  geom_text(aes(label = n, vjust = -.5)) + 
  theme(
    panel.background = element_rect(fill = "white", color = NA),
    axis.text.y = element_blank(), 
    axis.ticks.y = element_blank(),
    plot.title = element_text(hjust = 0.5)
  )
```

### 2.5 Loughran Lexicon

Here, I will use loughran lexicon instead of one of the lexicons used in the sample code.

```{r echo=TRUE, eval=TRUE}
lghrn <- get_sentiments("loughran")
unique(lghrn$sentiment)

#let’s explore the lexicon to see what types of words are litigious and constraining.

bondage_index %>% 
  unnest_tokens(word, text) %>% 
  inner_join(get_sentiments("loughran")) %>%
  filter(sentiment %in% c("litigious", "constraining")) %>%
  count(word, sentiment, sort = TRUE) %>%
  group_by(sentiment) %>%
  top_n(10) %>%
  ggplot() + 
  aes(x = reorder(word,desc(n)), y = n) + 
  geom_col() +
  facet_grid(~sentiment, scales = "free_x")  + 
  geom_text(aes(label = n, vjust = -.5)) + 
  labs(title = "Words Associated with Litigious and Constraining") + 
  ylab("Count") + 
  xlab("Word") + 
  theme(
    panel.background = element_rect(fill = "white", color = NA),
    axis.text.y = element_blank(), 
    axis.ticks.y = element_blank(),
    plot.title = element_text(hjust = 0.5)
  )
```

