---
title: "Data 607 Assignment 7"
author: "Trishita Nath"
date: "3/20/2021"
output: html_document
---

```{r echo=TRUE, eval=TRUE}
library(XML)
library(rjson)
library(httr)
library(tidyverse)
```
## Introduction

I prepared three files with the same data in the three specified formats: HTML, JSON and XML. These files are hosted on GitHub.   
I captured the following attributes of books:  
* Title
* Authors
* Publisher
* Year of Publication  
Each of the files was manually written and then loaded into Github for reference / reproducibility.  

## Data Loading

### XML Data

```{r echo=TRUE, eval=TRUE}
xml_data_raw <- xmlParse(GET("https://raw.githubusercontent.com/nathtrish334/Data-607/main/bookdata.xml"))
```
### HTML Data

```{r echo=TRUE, eval=TRUE}
html_data_raw <- readLines("https://raw.githubusercontent.com/nathtrish334/Data-607/main/bookdata.html")
# Pattern for the <td> html tags
regex_td = "<td>([0-9A-Za-z\\., \\(\\)-\\']*)</td>"
# Extract the lines that matching the pattern
regx_lines = grep(regex_td, html_data_raw[1:length(html_data_raw)], value=TRUE)
# Extract the required information from <td> tag
html_data_raw <- trimws(unlist(sub("<td>([0-9A-Za-z\\., \\(\\)-\\']*)</td>", "\\1", regx_lines)))
#str(html_data_raw)
```
### JSON Data

```{r echo=TRUE, eval=TRUE}
json_data_raw <- fromJSON(file="https://raw.githubusercontent.com/nathtrish334/Data-607/main/bookdata.json")
```

## Data Frames

### XML Data Frame

```{r echo=TRUE, eval=TRUE}
xml_data_df <- xmlToDataFrame(xml_data_raw)
xml_data_df
```

### HTML Data Frame

```{r echo=TRUE, eval=TRUE}
html_data_df <- cbind.data.frame(split(html_data_raw, rep(1:4, times=length(html_data_raw)/4)), stringsAsFactors=F)
names(html_data_df) <- c("Title", "Authors", "Publisher", "Year")
html_data_df
#str(html_data_df)
```

### JSON Data Frame

```{r echo=TRUE, eval=TRUE}
json_data_df <- data.frame(do.call("rbind", json_data_raw))
# Unnest data and convert it back to a df
json_data_df <- unnest(json_data_df, cols = c(Title, Authors, Publisher, Year)) %>% data.frame()
json_data_df
```

## Comparison of the Data Frames

```{r echo=TRUE, eval=TRUE}
library(diffobj)
identical(xml_data_df,html_data_df)
identical(xml_data_df,json_data_df)
identical(json_data_df,html_data_df)

#Check differences
diffObj(xml_data_df,json_data_df)
diffObj(json_data_df,html_data_df)
```

The only notable difference in the data frames is the 'Year' column. This is has a type of **chr** in html and xml data frames, whereas its type in the json data frame is **num**.