---
title: "Assignment9"
author: "Trishita Nath"
date: "3/30/2021"
output: 
  html_document:
    code_folding: show
    df_print: paged
    toc: true
    toc_float: false
    toc_collapsed: false
    smooth_scroll: false
    toc_depth: 3
number_sections: true
theme: paper
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment – Web APIs

The New York Times web site provides a rich set of APIs, as described here: https://developer.nytimes.com/apis
You’ll need to start by signing up for an API key.
Your task is to choose one of the New York Times APIs, construct an interface in R to read in the JSON data, and
transform it into an R DataFrame.  

## Introduction

I will make a call to the topstories API to read the JSON data and transform it into an R data frame. Specifically I will be working with the sports API: https://api.nytimes.com/svc/topstories/v2/sports.json?api-key=WSkYmmMA8AoszM0tsfyPRX33G3GNSSzJ  

My API key is _WSkYmmMA8AoszM0tsfyPRX33G3GNSSzJ_


## Loading the required packages

```{r}
library(httr)
library(jsonlite)
library(dplyr)
library(kableExtra)
```

## Reading and Transforming Data

```{r}
api_url <- "https://api.nytimes.com/svc/topstories/v2/sports.json?api-key=WSkYmmMA8AoszM0tsfyPRX33G3GNSSzJ"

# Read JSON data from the API
json_object <- fromJSON(api_url)

# Conver JSON to DataFrame
travel_df <- json_object$results
class(travel_df) # COnfirm it is a dataframe

# Column names
colnames(travel_df)

# Select relevant columns
travel_df <- subset(travel_df, select = c(section, title, abstract, url, byline))

# Show the first 50 records
head(travel_df,50)

```