---
title: "Data 607 Project 2"
author: "Trishita Nath"
date: "3/12/2021"
output: html_document
---

## Introduction
### Preparing different datasets for downstream analysis work.

Since I am a fan of movies, I am going to analyze the movies data-set. I will perform the following steps

* Deal with missing data 
* Proper naming of columns
* Perform visual analysis

My focus will be to analyze the gross and budget for Pirates of the Carribbean Films in the data set
```{r eval=TRUE}
library(readr)
library(tidyr)
library(tidyverse)
library(reshape2)
library(dplyr)
library(ggplot2)
```

```{r echo=TRUE, eval=TRUE}
movies_data <- read.csv("https://raw.githubusercontent.com/nathtrish334/Data-607/main/movie_dataset.csv",header=TRUE,sep=",")
head(movies_data[1:20])
```
```{r echo=TRUE, eval=TRUE}
nrow(movies_data) # Total records
names(movies_data) # Column names
colSums(movies_data == '' | is.na(movies_data)) # columns with blanks and NA
```
gross year has highest NA  and/or blank entries = 884

### Cleaning Data

Removing NA's
```{r echo=TRUE, eval=TRUE}
movies_data_clean<-na.omit(movies_data)
nrow(movies_data_clean)
```
movie title, budget, and gross for Pirates of the Carribbean films
```{r echo=TRUE, eval=TRUE}
pirates_data<-movies_data_clean[grep("Pirates of the Caribbean:", movies_data_clean$movie_title), ]
head(pirates_data)
```
Extract movie title, budget, and gross
```{r echo=TRUE, eval=TRUE}
pirates_data_cleaned<-subset(pirates_data, select=c(movie_title, budget, gross))
head(pirates_data_cleaned)
pirates_data_cleaned<-data.frame(pirates_data_cleaned)
```
Melt this subset using the reshape library
```{r echo=TRUE, eval=TRUE}
pirates_melted <- melt(pirates_data_cleaned, id = 'movie_title')
head(pirates_melted)
```
stacked bar plot for gross vs budget for each of Pirates of the Carribean Films
```{r echo=TRUE, eval=TRUE}
ggplot() + geom_bar(aes(y = value, x = movie_title, fill =variable ), data = pirates_melted,stat="identity")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  labs( x="Pirates Film", y="Amount in Dollars")
```

For each Pirates film, the gross was larger than the budget