---
title: "DATA 607 - Project 3"
author: "Trishita Nath"
date: "3/28/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## INTRODUCTION

I came across a study conducted by Jeff Hale in 2018 about *The Most In-Demand Skills for Data Scientists*. More can be found [here](https://towardsdatascience.com/the-most-in-demand-skills-for-data-scientists-4a4a8db896db)   
I will be analyzing his two datasets to determine ***“Which are the most valued data science skills?”***

## Data Source

I will use the the two datasets from John Hale's study. I have hosted them on my GitHub repository. 

```{r echo=TRUE, eval=TRUE}
gen_skills_url <- "https://raw.githubusercontent.com/nathtrish334/Data-607/main/ds_general_skills_revised.csv"
soft_skills_url <- "https://raw.githubusercontent.com/nathtrish334/Data-607/main/ds_job_listing_software.csv"
```

## DATA LOADING

```{r echo=TRUE, eval=TRUE}
library(tidyverse)
library(ggplot2)
library(rvest)
library(tm)
library(SnowballC)
library(wordcloud2)
library(stringr)
library(plotly)
```

```{r echo=TRUE, eval=TRUE}
gen_skills <- read_csv(file=gen_skills_url)
soft_skills <- read_csv(file=soft_skills_url)
```

## Data Cleaning  
I cleaned the two datasets and joined them into one

```{r echo=TRUE, eval=TRUE}
df_1 <- as_tibble(gen_skills, key = "ï..Keyword")
df_1$Skill_Set <- "Soft"
df_2 <- as_tibble(soft_skills, key = "ï..Keyword")
df_2$Skill_Set <- "Tech"
df_2 <- select(df_2,-c(6,7,8,9,10,11,12))
combined_df<-rbind(df_1, df_2)
head(combined_df)

# Rename the variable
names(combined_df)[names(combined_df) == "ï..Keyword"] <- "Keyword"

#Tidying the combined dataframe
combined_df$Total <- rowSums(sapply(combined_df[,c(2:5)], as.numeric))
combined_df <- mutate(combined_df[, c(1,6,2:5,7)])
combined_df <- data.frame(combined_df)
colnames(combined_df)
```

Averages of keywords

```{r echo=TRUE, eval=TRUE}
SimplyHired <- 3829
Monster <- 3746
Indeed <- 5138
LinkedIn <- 8610
Total <- LinkedIn + Indeed + SimplyHired + Monster
combined_df$LinkedIn <- ((combined_df$LinkedIn)/(LinkedIn))*100
combined_df$Indeed <- ((combined_df$Indeed)/(Indeed))*100
combined_df$SimplyHired <- ((combined_df$SimplyHired)/(SimplyHired))*100
combined_df$Monster <- ((combined_df$Monster)/((Monster)))*100
combined_df$Total <- ((combined_df$Total/((Total))))*100
dim(combined_df)
head(combined_df, 10)
```

## ANALYSIS

### Top 10 Soft Skills for Data Scientists
```{r echo=TRUE, eval=TRUE}
combined_df_soft <- filter(combined_df, combined_df$Skill_Set == "Soft")
dim(combined_df_soft)

combined_df_soft_top10 <- arrange(combined_df_soft[1:10,c(1,3:7)], desc(Total))
combined_df_soft_top10

# Top ten Soft skills as a percentage of total Data Science jobs

plot_ly(combined_df_soft_top10, 
        x = ~ combined_df_soft_top10, 
        y = combined_df_soft_top10$Total, 
        type = 'bar',
        name = combined_df_soft_top10$Keyword)%>% 
layout(title = 'Top 10 Soft Skills for Data Scientists',yaxis = list(title = '% of Data Science Jobs'), xaxis = list(title = 'Soft Skills'))
```

### Top 10 Soft skills as a pecentage of total data science jobs by job site
```{r echo=TRUE, eval=TRUE}
soft_top10_by_site <- data.frame("Keyword"=combined_df_soft_top10, combined_df_soft_top10)

soft_visual_top10 <- soft_top10_by_site[,c('Keyword', 'LinkedIn', 'Indeed', 'SimplyHired', 'Monster', 'Total')]
head(soft_visual_top10, 5)

plot_ly(soft_visual_top10, x = ~Keyword, y = ~LinkedIn, type = 'bar', name = "Linkedin") %>%
add_trace(y = ~Indeed, name = "Indeed") %>% 
add_trace(y = ~SimplyHired, name = "SimplyHired") %>% 
add_trace(y = ~Monster, name = "Monster") %>% 
add_trace(y = ~Total, name = "Average") %>% 
layout(title = 'Soft Skills by Job Sites', yaxis = list(title = '% of Data Science Jobs'), xaxis = list(title = 'Soft Skills'), barmode = 'group')
```

### Top 10 Technical Skills for Data Scientists
```{r echo=TRUE, eval=TRUE}
combined_df_tech <- filter(combined_df, combined_df$Skill_Set == "Tech")
dim(combined_df_tech)

combined_df_tech_top10 <- arrange(combined_df_tech[1:10,c(1,3:7)], desc(Total))
combined_df_tech_top10

# Top 10 Technical skills as a percentage of total Data Science jobs

plot_ly(combined_df_tech_top10, 
        x = ~ combined_df_tech_top10, 
        y = combined_df_tech_top10$Total, 
        type = 'bar', 
        name = combined_df_tech_top10$Keyword)%>% 
layout(title = 'Top 10 Technical Skills for Data Scientists',yaxis = list(title = '% of Data Science Jobs'), xaxis = list(title = 'Technical Skills'))
```

### Top 10 Technical skills as a pecentage of total data science jobs by job site
```{r echo=TRUE, eval=TRUE}
tech_top10_by_site <- data.frame("Keyword"=combined_df_tech_top10, combined_df_tech_top10)

tech_visual_top10 <- tech_top10_by_site[,c('Keyword', 'LinkedIn', 'Indeed', 'SimplyHired', 'Monster', 'Total')]
head(tech_visual_top10, 5)

plot_ly(tech_visual_top10, x = ~Keyword, y = ~LinkedIn, type = 'bar', name = "Linkedin") %>%
add_trace(y = ~Indeed, name = "Indeed") %>% 
add_trace(y = ~SimplyHired, name = "SimplyHired") %>% 
add_trace(y = ~Monster, name = "Monster") %>% 
add_trace(y = ~Total, name = "Average") %>% 
layout(title = 'Technical Skills by Job Sites', yaxis = list(title = '% of Data Science Jobs'), xaxis = list(title = 'Technical Skills'), barmode = 'group')
```

### Top Soft and Technical skills needed for a data scientist
```{r echo=TRUE, eval=TRUE}
# Top Soft skill required
head(combined_df_tech,1)

# Top Technical skill required
head(combined_df_soft,1)
```

## CONCLUSION

According to this dataset, the top soft skill that employers are looking for when hiring a data scientist is Analysis. The top technical skill being targeted is Python.